Response to reviewers

### 2025-08-09 src/docs_paper-jtb/paper.Rmd
Update text regarding the NUTS diagnostics

\begin{table}[]
\small
\begin{tabularx}{\textwidth}{p{0.15\linewidth}p{0.15\linewidth}p{0.6\linewidth}}
\toprule
Method & Software & Details \\
\midrule
GEB & \texttt{TMB} & $1000$ samples \\
GPCA-AGHQ & \texttt{aghq} & $k = 3, s = 8$, 1000 samples \\
NUTS & \texttt{tmbstan} & 4 chains of 100000 iterations, with the first half of each chain discarded as warmup, thinned by a factor of 20, to give a total of 10000 samples. Default NUTS tuning parameters \citep{hoffman2014no}. \\
\bottomrule
\end{tabularx}
\caption{A summary of settings used for each inferential method.}
\label{tab:inference-methods}
\small
\end{table}

@@

The Naomi model was difficult to efficiently sample from using NUTS via `tmbstan`.
The lowest effective sample size (ESS) was `r round(naomi_nuts$ess_min, 0)` (2.5% quantile `r round(naomi_nuts$ess_lower, 0)`, 50% quantile `r round(naomi_nuts$ess_median, 0)`, and 97.5% quantile `r round(naomi_nuts$ess_upper, 0)`; Supplementary Figure 1A).
The largest potential scale reduction factor ($\hat R$) was `r round(naomi_nuts$rhat_max, 3)` (2.5% quantile `r round(naomi_nuts$rhat_lower, 3)`, 50% quantile `r round(naomi_nuts$rhat_median, 3)`, and 97.5% quantile `r round(naomi_nuts$rhat_upper, 3)`; Supplementary Figure 2).
Supplementary Figure 3 shows traceplots for the parameters with the lowest ESS and $\hat R$ respectively.
Though inaccuracies remain possible, these diagnostics are sufficient to treat inferences obtained from NUTS as a gold-standard.

>>>>

\begin{table}[]
\small
\begin{tabularx}{\textwidth}{p{0.15\linewidth}p{0.15\linewidth}p{0.6\linewidth}}
\toprule
Method & Software & Details \\
\midrule
GEB & \texttt{TMB} & $1000$ samples \\
GPCA-AGHQ & \texttt{aghq} & $k = 3, s = 8$, 1000 samples \\
NUTS & \texttt{tmbstan} & 18 chains of 40000 iterations, with the first half of each chain discarded as warmup, thinned by a factor of 20, to give a total of 18000 samples. The target average acceptance probability was set to 0.95 and the maximum treedepth to 12 \citep{hoffman2014no}. \\
\bottomrule
\end{tabularx}
\caption{A summary of settings used for each inferential method.}
\label{tab:inference-methods}
\small
\end{table}

@@

The Naomi model was challenging to efficiently sample from using NUTS via `tmbstan`.
We used 18 chains of 40000 iterations run on a high-performance computing cluster (Table \@ref(tab:inference-methods)).
The lowest resulting effective sample size (ESS) was `r round(naomi_nuts$ess_min, 0)` (2.5% quantile `r round(naomi_nuts$ess_lower, 0)`, 50% quantile `r round(naomi_nuts$ess_median, 0)`, and 97.5% quantile `r round(naomi_nuts$ess_upper, 0)`; Supplementary Figure 1A).
The largest resulting potential scale reduction factor ($\hat R$) was `r round(naomi_nuts$rhat_max, 3)` (2.5% quantile `r round(naomi_nuts$rhat_lower, 3)`, 50% quantile `r round(naomi_nuts$rhat_median, 3)`, and 97.5% quantile `r round(naomi_nuts$rhat_upper, 3)`; Supplementary Figure 2).
Supplementary Figure 3 shows traceplots for the parameters with the lowest ESS and $\hat R$ respectively.
Based on these diagnostics, we considered the NUTS samples to be of a high quality and suitable for use as a gold-standard.

### 2025-08-09 src/docs_paper-jtb/revisions.Rmd
Update text regarding the NUTS diagnostics

We agree that running more iterations of the NUTS sampler is a good idea, and could improve the minimum ESS and accuracy of any comparisons, especially distributional ones.
We have rerun the NUTS sampler using a HPC with the 18 chains (across 18 cores) with 120,000 iterations each.

>>>>

We agree that running more iterations of the NUTS sampler is a good idea, and could improve the minimum ESS and accuracy of any comparisons, especially distributional ones.
We have rerun the NUTS sampler using a HPC with 18 chains (across 18 cores) with 40,000 per chain and a thinning factor of 20.
To improve the sampling also increased the `adapt_delta` parameter from the default of 0.8 to 0.95, and the `max_treedepth` parameter from the default of 10 to 12.
The resulting diagnostics are much improved, as described in an update to Section 5.1:

```{r}
naomi_nuts <- readRDS("depends/mcmc-out.rds")
```

> The Naomi model was challenging to efficiently sample from using NUTS via `tmbstan`.
We used 18 chains of 40000 iterations run on a high-performance computing cluster (Table \@ref(tab:inference-methods)).
The lowest resulting effective sample size (ESS) was `r round(naomi_nuts$ess_min, 0)` (2.5% quantile `r round(naomi_nuts$ess_lower, 0)`, 50% quantile `r round(naomi_nuts$ess_median, 0)`, and 97.5% quantile `r round(naomi_nuts$ess_upper, 0)`; Supplementary Figure 1A).
The largest resulting potential scale reduction factor ($\hat R$) was `r round(naomi_nuts$rhat_max, 3)` (2.5% quantile `r round(naomi_nuts$rhat_lower, 3)`, 50% quantile `r round(naomi_nuts$rhat_median, 3)`, and 97.5% quantile `r round(naomi_nuts$rhat_upper, 3)`; Supplementary Figure 2).
Supplementary Figure 3 shows traceplots for the parameters with the lowest ESS and $\hat R$ respectively.
Based on these diagnostics, we considered the NUTS samples to be of a high quality and suitable for use as a gold-standard.
